Evolution of Processes in Delivering Reliable Software within the European Scientific Arena

### 2018-09-25 paper-ieee-trans.tex
Remove bugzilla figure and rephrasing

@@

The {\sl EGI--InSPIRE} (Integrated Sustainable Pan--European Infrastructure for
Researchers in Europe, May 2010 -- May 2014) project \cite{cordis:egi-inspire}
was the continuation of the {\sl EGEE--III} project, with the objective of establishing and
maintaining a sustainable European Grid Infrastructure, composed by a federation of National
Grid Initiatives and interoperable with other Grids worldwide.
This goal would be accomplished through the
development and maintenance of various operational tools -- such as the Operations Portal \cite{egi-ops},
the {\sl EGI} Helpdesk \cite{ggus}, or the Grid configuration database (GOCDB) \cite{gocdb} -- and the
management of the software provisioning process \cite{mario}, dealing with the validation and distribution of the software to
the production infrastructures. As a result, this process led to a production--ready
Grid computing middleware distribution named {\sl UMD} (Unified Middleware Distribution). At that time the {\sl UMD}
was a stack of about 250 software components from several technology providers, such
as {\sl EMI} and {\sl Globus}. In order to be distributed through UMD
repositories, the software had to be compliant with EGI's Quality Criteria (QC) \cite{egi-qc}. The
QC defined a set of requirements in different areas: documentation, deployment, security, information
model and operations.
{\sl UMD} is still being used and
deployed in the European scientific e--Infrastructures under the follow--up
project {\sl EGI--Engage} (Engaging the EGI Community towards an Open Science
Commons) \cite{cordis:egi-engage}. This distribution is currently complemented
by a Cloud--specific one called {\sl CMD} (Cloud Middleware Distribution).

>>>>

Project monitoring and reporting mechanisms allowed to assess the risks, efforts and documentation
produced. Development guidelines \cite{datagridguide} were produced, covering the various phases, including
1) packaging, 2) test and validation, and 3) style and naming convention. The specification of quality indicators
\cite{datagridindicators} facilitated the detection of deviations in the collected metric data. Figure 
\ref{fig:efficiency} shows the efficiency performance indicator, measured in production environments as the
number of jobs successfully completed divided by the total number of jobs submitted. As it can be observed,
there is a positive trend of efficiency over the months, with the unique exception of the LCG1 experiment.
\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth, height=50mm]{images/dgeff.png}
\caption{Efficiency \% = Number of jobs successfully completed / Total Number of jobs submitted: results are provided for five experiments involved in the project.}
\label{fig:efficiency}
\end{figure}

@@

The {\sl EGI--InSPIRE} (Integrated Sustainable Pan--European Infrastructure for
Researchers in Europe, May 2010 -- May 2014) project \cite{cordis:egi-inspire}
was the continuation of the {\sl EGEE--III} project, with the objective of establishing and
maintaining a sustainable European Grid Infrastructure, composed by a federation of National
Grid Initiatives and interoperable with other Grids worldwide.
This goal would be accomplished through the
development and maintenance of various operational tools -- such as the Operations Portal \cite{egi-ops},
the {\sl EGI} Helpdesk \cite{ggus}, or the Grid configuration database (GOCDB) \cite{gocdb} -- and the
management of the software provisioning process \cite{mario}, dealing with the validation and distribution of the software to
the production infrastructures. As a result, this process led to a production--ready
Grid computing middleware distribution named {\sl UMD} (Unified Middleware Distribution). At that time the {\sl UMD}
was a stack of about 250 software components from several technology providers, such
as {\sl EMI} and {\sl Globus}. In order to be distributed through UMD
repositories, the software had to be compliant with EGI's Quality Criteria (QC) \cite{egi-qc}. The
QC defined a set of requirements in different areas: documentation, deployment, security, information
model and operations.

The impact of the quality assurance activities was observed once the EMI software contributed to the 
EGI--InSPIRE project. The data reported in Table \ref{tab:emi} (see page 13 in \cite{emi-quality-model})
summarizes the EMI software quality evolution per project quarters (PQ) as evaluated by the EGI project
with the help of the UMD QC. We can observe the improvement in quality of the EMI releases over time, 
measured by the number of products that both met the software distribution criteria, by means of the UMD QC
definition, and the EGI Stage Rollout (SR) phase. This latter phase completed the criteria validation by 
deploying the UMD QC--certified software in a set of candidate production sites.

\begin{table}[!h]
\renewcommand{\arraystretch}{1.3}
\caption{EMI software quality evolution}
\label{tab:emi}
\centering
\begin{tabular}{lp{1cm}p{1cm}p{1cm}p{1cm}}
\hline
\hline
PQ & Release Products (RP) & N. RP Passed UMD QC & N. RP Passed SR & N. RP Failed UMD QC\\
\hline
\hline
5 & 30 & 27 & 27 & 0\\
6 & 30 & 28 & 26 & 2 \\
7 & 27 & 26& 24& 2\\
8 & 18 & 18 & 18 & 0\\
\hline
\hline
\end{tabular}
\end{table}

{\sl UMD} is still being used and
deployed in the European scientific e--Infrastructures under the follow--up
project {\sl EGI--Engage} (Engaging the EGI Community towards an Open Science
Commons) \cite{cordis:egi-engage}. This distribution is currently complemented
by a Cloud--specific one called {\sl CMD} (Cloud Middleware Distribution).
